Designing and modeling a browser-based distributed evolutionary computation systems

### 2015-03-31 modeling-volunteer-computing-gecco.tex
Advancing introduction

\begin{abstract}
Browsers have scaled from simple page-rendering engines to operating
systems that include most services the lower OS layer has, with the
added facility that applications can be run by just visiting a web
page. In this paper we will describe the front and back end of a
distributed evolutionary computation system that uses the browser's
capabilities of running programs written in JavaScript. We will focus on two
different aspects of volunteer computing: first, the
pragmatic: where to find those resources, which ones can be used, what
kind of support you have to give them; and then, the theoretical: how
evolutionary algorithms can be adapted to an environment in which nodes
come and go, have different computing capabilities and operate in
complete asynchrony of each other. We will examine the setup needed to
create a  simple distributed evolutionary algorithm using
JavaScript and then find a model of how users react to it by
collecting data from several experiments featuring different classical
benchmark functions. 
\end{abstract}

@@

In general, it should be a requisite for volunteer computing
experiment will have to be performed 
``in the open''. The fact that somebody is giving to you, basically for
free, computing resources, means that the scientist using them has to
give back. The baseline is releasing the source code used: all
volunteer computing platforms, from SETI@home on, have done it. The
opposite is probably the reason why many companies like PopularPower \cite{buyya2001compute}
have folded or others like CrowdProcess have decided to turn their
product to in-house computation: there must be a mutual relationship
of trust among the scientist and the person that is running his/her
code in the browser. As it has been mentioned in the early stages of what
was then called {\em desktop grid computing} \cite{gc:bausch}, in
fact CPU cycle selling might not make economic sense since there are
not so much demand for it and supply is quite high. However, while
potential supply is in fact huge, {\em actual} supply will depend on
the person holding it willing to actually allocate it to a particular
company selling it or a particular experiment needing it, not to
mention the fact that the experiment {\em actually} has to draw the
attention of the supplier. That is why trust is essential, and using
free software might not be enough: Openness
has to progress from open source code to open science: releasing all
data obtained in the experiment in a repository such as GitHub,
mentioned above, and even allowing real-time access to experimental data
to users.

% Había metido en el .bib el artículo de detectar vulnerabilidades 
% en EAs distribuidos de Juanlu, pero creo que al final no tiene sentido 
% hablar en este paper de él (Muszynski14vulnerabilities)

@@

The server was also written using node.js and the {\tt express.js}
module and has three REST \cite{Castillo12REST} %FERGU: cito a este paper que es un capítulo de libro en vez de Arxiv
functions, that is, three functions that can be accessed from any
client, including the browser, via a request encoded in an URL and a
standard HTTP command. These functions are\begin{itemize}
\item {\tt GET random} returns a random, non-evaluated individual,
  from the pool. The pool is just an array containing chromosomes,
  that is initialized at the beginning of the experiment.
\item {\tt PUT one} puts a single individual in the pool. In REST
  conventions, {\tt PUT} is used to create a resource, that is why it
  is used, in this case, to add a new individual to the pool.
\item {\tt GET log} returns the experiment log so far. It is used by
  the researcher to gather data, and as a transparency measure so that
  anyone can use it. 
\end{itemize}

>>>>

\begin{abstract}
Web browsers have scaled from simple page-rendering engines to operating
systems that include most services the lower OS layer has, with the
added facility that applications can be run by just visiting a web
page. In this paper we will describe the front and back end of a
distributed evolutionary computation system that uses the browser's
capabilities of running programs written in JavaScript. We will focus on two
different aspects of volunteer computing: first, the
pragmatic: where to find those resources, which ones can be used, what
kind of support you have to give them; and then, the theoretical: how
evolutionary algorithms can be adapted to an environment in which nodes
come and go, have different computing capabilities and operate in
complete asynchrony of each other. We will examine the setup needed to
create a  simple distributed evolutionary algorithm using
JavaScript and then find a model of how users react to it by
collecting data from several experiments featuring different classical
benchmark functions. 
\end{abstract}

@@

In general, it should be a requisite for volunteer computing
experiments  to be performed 
``in the open''. The fact that somebody is giving to you, basically for
free, computing resources, means that the scientist using them has to
give back; besides, people do not usually visit a resource unless they
can trust it; trust is an essential factor in the sharing of resources
such as this one, computing power \cite{van2009would}. The point of
departure for establishing this trust is releasing the source code used: all
volunteer computing platforms, from SETI@home on, have done it. The
opposite is probably the reason why many companies like PopularPower \cite{buyya2001compute}
have folded or others like CrowdProcess have decided to turn their
product to in-house computation \cite{ycombinator}: there must be a mutual relationship
of trust among the scientist and the person that is running his/her
code in the browser. As it has been mentioned in the early stages of what
was then called {\em desktop grid computing} \cite{gc:bausch}, in
fact CPU cycle selling might not make economic sense since there are
not so much demand for it and supply is quite high. However, while
potential supply is in fact huge, {\em actual} supply will depend on
the person holding it willing to actually allocate it to a particular
company selling it or a particular experiment needing it, not to
mention the fact that the experiment {\em actually} has to draw the
attention of the supplier, and thus it is better analyzed using the
Uses and Gratifications Theory \cite{severin2010communication}, that
is, taking into account {\em why} an user would do it, than using a
purely computational approach. That is why considering trust is
essential, and using free software might not be enough: Openness
has to progress from open source code to open science: releasing all
data obtained in the experiment in a repository such as GitHub,
mentioned above, and even allowing real-time access to experimental data
to users.

% Había metido en el .bib el artículo de detectar vulnerabilidades 
% en EAs distribuidos de Juanlu, pero creo que al final no tiene sentido 
% hablar en este paper de él (Muszynski14vulnerabilities)
%Lo voy a meter
In fact, this relationship of trust is in many cases the only
mechanism preventing the malicious use of the system by creating bogus
solutions or simply attempting to exploit vulnerabilities by injecting
code. These problems have been analyzed in
\cite{Muszynski14vulnerabilities} but in principle our working
hypothesis will be that the clients using the browser are not
manipulating the code make the experiment finish before it should,
which is possibly the only way in which the experiment would be
affected.

@@

The server was also written using node.js and the {\tt express.js}
module and has three REST \cite{Castillo12REST} %FERGU: cito a este paper que es un capítulo de libro en vez de Arxiv
functions, that is, three functions that can be accessed from any
client, including the browser, via a request encoded in an URL and a
standard HTTP command. These functions are\begin{itemize}
\item {\tt GET random} returns a random, non-evaluated individual,
  from the pool. The pool is just an array containing chromosomes,
  that is initialized at the beginning of the experiment.
\item {\tt PUT one} puts a single individual along with its fitness in the pool. In REST
  conventions, {\tt PUT} is used to create a resource, that is why it
  is used, in this case, to add a new individual to the pool.
\end{itemize}
